A comparison of two second order traffic flow models

### 2015-03-13 FirstDraftMaterials/Draft.tex
Adding code to show approximating examples, and general edits to draft.

The LWR model uses the flux function $f(\rho) = \rho V(\rho)$, where $V(\rho)$ represents the preferred, or equilibrium, velocity, a given nonincreasing function of $\rho$, nonnegative for $\rho$ between 0 and $\rho_m$, the ``jam" density.  The LWR model behaves well macroscopically, predicting piece-wise smooth density with transitions between regions approximated by shocks.  However, it assumes that cars change their velocities instantaneously as they travel through shocks.  One way to eliminate shocks would be to add a diffusion term: $\rho_t + \Big(\rho V(\rho))_x = \mu \rho_{xx}$.  This additional term is meant to account for the driver's awareness of the road ahead and behind.  However, this model can be shown to predict negative velocities in some instances.

@@

Restrictions on the way that $\lambda$ must vary across a wave allow us to identify the regions of the Hugoniot loci and integral curves that are valid for any given point. When traveling across a shock from a left state, $q_l$, to a right state $q_r$, $\lambda$ must decrease. When traveling across a rarefaction wave from a left state, $q_l$, to a right state $q_r$, $\lambda$ must increase. In Figure \ref{fig:PW_curves} (c) and (d) we see the Hugoniot loci and integral curves for this system with the contour lines for $\lambda_1$ and $\lambda_2$. Using these plots, we can determine which areas of the curves are valid shocks and which are valid rarefaction waves. Figure \ref{fig:PW_curves} (e) and (d) shows the points $q_l$ and $q_r$, respectively, with only the valid waves displayed. Here, shocks are shown in red or magenta and rarefaction waves are shown in either light or dark green.

@@

In conservative form, the second equation becomes $\Big( \rho (v + p(\rho)) \Big)_t + \Big( \rho v (v + p(\rho)) \Big)_x = 0$.  Aw and Rascle call this conserved quantity, $y = \rho v + \rho p(\rho)$, "momentum", although they admit that $y$ has no obvious physical interpretation. Furthermore, the second equation can also be rewritten as an equation for velocity: $v_t + \Big(v - \rho p'(\rho)\Big)v_x = 0$. If we consider the linearized system,

@@

\subsubsection{Hugoniot Loci}
In order to consider the Hugoniot loci for this system we must first write the system in conservation form. 
To accomplish this, note that multiplying the first equation by $(v + p(\rho ))$ and the second equation 
by $\rho$ gives us the two equations 
\begin{align*}
&(v + p(\rho ))\partial_t\rho + (v + p(\rho ))\partial_x(\rho v) = 0,  \\
&\rho\partial_t \left(v + p(\rho )\right) + \rho v\partial_x \left( v + p(\rho )\right) = 0.
\end{align*}
Adding these two equations gives
\begin{align}\label{AR:eq2}
\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
\end{align}
Then from (\ref{AR:eq1}) and (\ref{AR:eq2}) we have the system 
\begin{align*}
&\partial_t\rho + \partial_x(\rho v) = 0, \\
&\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
\end{align*}
This can be rewritten as $q_t + f(q)_x = $ by defining
\begin{align*}
q = \left[ \begin{matrix}
\rho \\ y
\end{matrix}\right], \hspace{0.3in}
f(q) = \left[ \begin{matrix}
v\rho \\
vy
\end{matrix}\right].
\end{align*}
where $y = \rho\left(v + p(\rho )\right)$.
The Rankine-Hugoniot condition tells us that 
\begin{align*}
s(q_*- q) = f(q_*) - f(q).
\end{align*}
Therefore, from this condition we get the two equations
\begin{align}
s(\rho_* - \rho) = v_*\rho_* - v\rho, \label{AR:eq3}\\
s(y_* - y) = v_*y_* - vy.\label{AR:eq4}
\end{align}
With a little rearranging and by plugging in for $y$, these two equations can be rewritten as
\begin{align}
&\rho_*(s - v_*) = \rho (s - v), \label{AR:eq5}\\
&\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) = \rho(s - v)\left(v + p(\rho )\right).\label{AR:eq6}
\end{align}
From (\ref{AR:eq5}), we know that there are two cases:

(a) $\rho_*(s - v_*) = \rho (s - v) = 0$. In this case we have that
\begin{align*}
s - v_* = s - v 
\end{align*}
except if one of the two densities, $\rho_*$, $\rho$ is zero. Therefore, we can say that
\begin{align*}
v_* = v.
\end{align*}
Since the velocity does not change, this is a contact discontinuity. Using the fact that $v = v_*$ in (\ref{AR:eq4}) gives
\begin{align*}
y_*(s - v) = y(s - v).
\end{align*}
So, either $y = y_*$ or $s = v$. If $y = y_*$ we have a stationary point in the $Y = (\rho, y)$ plane, which does not 
describe a wave. Therefore, $s = v$. Note that we have no restrictions on $y$ other than its definition. 
Therefore this loci in the $Y$ plane is given by 
\begin{align*}
y = \rho ( v_* + p(\rho )).
\end{align*}

@@

(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into (\ref{AR:eq6}) gives
\begin{align*}
\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
\rho(s - v)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
v_* + p(\rho_* )&= v + p(\rho ).
\end{align*}
Therefore, the Hugoniot loci for this case in the Y plane are given by
\begin{align*}
y = v_* + p(\rho_* ),
\end{align*}
since $y = v + p(\rho )$ by definition. Note that solving (\ref{AR:eq3}) for $s$ gives
\begin{align*}
s = \frac{\rho_*v_* - \rho v}{\rho_* - \rho}.
\end{align*}
Using the fact that $v_* = v + p(\rho) - p(\rho_*)$, we get that
\begin{align*}
s = \frac{\rho_*\left( v + p(\rho) - p(\rho_*)\right)- \rho v}{\rho_* - \rho}
= v - \rho_*\frac{p(\rho ) - p(\rho_*)}{\rho - \rho_*}
\approx v - \rho p'(\rho)
\end{align*}
for $\rho_* \approx \rho$.

In the $U = (\rho, v)$ plane these loci become (a) $v = v_*$, $\rho$ free, and (b) $v = v_* + p(\rho_* ) - p(\rho )$.
 In the $M = (\rho, \rho v)$ plane these loci become (a) $\rho v = \rho_* v_*$, $\rho$ free, 
 and (b) $\rho v = \rho \left( v_* + p(\rho_*)\right) - \rho p(\rho))$. Figure \ref{fig:AR_curves} shows the Hugoniot loci 
 in the different planes. TODO: ADD PARAMETERS USED FOR FIGURES? OR, WILL INCLUDING THE CODE MAKE THAT REDUNDANT? 
 IF WE ADD PARAMETERS, MAYBE THE BEST PRESENTATION WOULD BE IN A TABLE?

\subsubsection{Integral Curves}
In order to consider the integral curves for this system \cite{AwRascle2000} 
works from the system
\begin{align*}
&\partial_t \rho + \partial_x (\rho v) = 0 \\ 
&\partial_t v + \left(v - p'(\rho)\rho\right)\partial_x v = 0,
\end{align*}
where the second equation is found by multiplying (\ref{AR:eq1}) 
by $p'(\rho)$ and subtracting it from (\ref{AR:eq1.5}). 
This can be rewritten as $q_t + Aq_t = 0$ by defining
\begin{align*}
q = \left[ \begin{matrix}
\rho \\ v
\end{matrix}\right] , \hspace{0.3in}
A = \left[ \begin{matrix}
v & \rho \\
0 & v - p'(\rho ) \rho
\end{matrix}\right].
\end{align*}
The eigenvalues of $A$ are $\lambda_1 = v - p'(\rho ) \rho$ and 
$\lambda_2 = v$. Note that since $p(\rho ) = \rho^{\gamma}$ 
and $\rho \geq 0$, $\lambda_1 < \lambda_2$ so long as $\rho \neq 0$. 
Also note that if $\rho = 0$ then $A$ is no longer 
hyperbolic. The eigenvectors of $A$ are
\begin{align*}
r_1 = \left[ \begin{matrix}
1 \\ - p'(\rho )
\end{matrix}\right], \hspace{0.3in}
r_2 = \left[ \begin{matrix}
1 \\ 0
\end{matrix}\right].
\end{align*}
Using equation (13.24) from \cite{LeVeque2002} with $\alpha = 1$ tells us 
that the integral curves can be found by considering $q'(\xi ) = r_p$. 
For $p = 1$ this gives us the two equations
\begin{align*}
&\rho'(\xi ) = 1 \\
&v'(\xi ) = - p'(\rho ),
\end{align*}
which can be solved to find that we have parameratized our curve by $\rho$, 
and $v = - p'(\rho ) + c_1$ while $\rho$ is free, where $c_1$ is a constant. 
Since for a given point $(\rho_*, v_*)$ when $\rho = \rho_*$ we want 
$v = v_*$, we get that $c_1 = v_* + p'(\rho_*)$. Thus, the integral curve 
is given by 
\begin{align*}
v = v_* + p'(\rho_*) - p'(\rho ).
\end{align*}
This coincides with the Hugoniot loci from case (b). For $p = 2$, we get the 
two equations
\begin{align*}
&\rho'(\xi ) = 1 \\
&v'(\xi ) = 0,
\end{align*}
which can be solved to find that we have parameratized our curve by $\rho$, 
and that $v$ is a constant. Therefore, $v = v_*$ and $\rho$ is free. 
This coincides with the Hugoniot loci from case (a).

\begin{figure}[H]
 \centering
    \subfigure[$\lambda_1$-curves with contours of $\lambda_1$.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M.jpg}
   }
 \subfigure[Valid waves for the point $q_l$.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_ql.jpg}
   }
 \subfigure[Valid waves for the point $q_r$.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_qr.jpg}
   }
 \caption[Optional caption for list of figures]
 {Figures showing the calculation of the valid waves for a given point. Figures (a)
  and (d) show the $\lambda_1$ curves in black with the $\lambda_1$ contours 
  in cyan. The other figures show valid shocks in red and valid rarefaction 
  waves in green. The contact discontinuity is shown in blue.}
   \label{fig:AR_validity}
\end{figure}

@@

Therefore, it is important to take into account the contour lines of $\lambda$ when 
looking at the loci and integral curves. Figures \ref{fig:AR_validity}(a) and 
\ref{fig:AR_validity}(d) show the $\lambda_1$-curves in the $U$ and $M$ planes,
with the contour lines of $\lambda_1$ shown in magenta. This information is used
to determine which areas of the $\lambda_1$ curves are valid shock waves and which
are valid rarefaction waves. The rest of the subfigures show the regions of validity for 
specific points $q_*$, depending on whether $q_* = q_l$ or $q_* = q_r$. For these plots,
$q_* = (\rho_*,v_*) = (0.2, 0.2)$. Note that the contact discontinuity is valid
anywhere in the plane.

@@

\begin{figure}[H]
 \centering
 \subfigure[Solution from the AR model.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example1.jpg}
   }
    \subfigure[Solution from the PW model.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example1.jpg}
   }
 \caption[Optional caption for list of figures]
 {Solutions to the Riemann problem for example 1 from the AR and PW models.}
  \label{fig:example1}
\end{figure}

\begin{figure}[H]
 \centering
 \subfigure[Solution from the AR model.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example2.jpg}
   }
 \subfigure[Solution from the PW model.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example2.jpg}
   }
 \caption[Optional caption for list of figures]
 {Solutions to the Riemann problem for example 2 from the AR and PW models.}
  \label{fig:example2}
\end{figure}

@@

Reproducing the Riemann problems found in \cite{AwRascle2000}, we consider two different examples. The variables used in these two examples are shown in Table \ref{table:1}. Note that in the solution from the AR model for example 1, shown in Figure \ref{fig:example1} (a), the transition from $q_l = (\rho_l, v_l\rho_l)$ to $q_r = (\rho_r, v_r\rho_r)$ is first a shock and then a contact discontinuity. The middle state is denoted by $q_m$ in the figure. In the solution given by the AR model for example 2, show in Figure \ref{fig:example2} (a), the transition from $q_l = (\rho_l, v_l\rho_l)$ to $q_r = (\rho_r, v_r\rho_r)$ is a shock wave, and there is no middle state.

In both examples the density and velocity are greater in the right state than in the left state.  This corresponds to the situation considered by Aw and Rascle that motivated the use of the convective derivative.  In the first example, we see that the AR model predicts that the density of the middle state is less than both left or right states, but that the velocity increases steadily from left state to right.  This is what we would expect, as drivers spread out as the accelerate to catch up with the cars ahead.  On the other hand, the PW model predicts that the middle state will have a greater density than the left state but a smaller velocity.  This is opposite of what we would expect.

@@

>>>>

The LWR model uses the flux function $f(\rho) = \rho V(\rho)$, where $V(\rho)$ represents the preferred, or equilibrium, velocity, a given nonincreasing function of $\rho$, nonnegative for $\rho$ between 0 and $\rho_m$, the ``jam" density.  The LWR model behaves well macroscopically, predicting piece-wise smooth density with transitions between regions approximated by shocks.  However, it assumes that cars change their velocities instantaneously as they travel through shocks.  One way to eliminate shocks would be to add a diffusion term: $\rho_t + \Big(\rho V(\rho)\Big)_x = \mu \rho_{xx}$.  This additional term is meant to account for the driver's awareness of the road ahead and behind.  However, this model can be shown to predict negative velocities in some instances.

@@

Restrictions on the way that $\lambda$ must vary across a wave allow us to identify the regions of the Hugoniot loci and integral curves that are valid for any given point. When traveling across a shock from a left state, $q_l$, to a right state $q_r$, $\lambda$ must decrease. When traveling across a rarefaction wave from a left state, $q_l$, to a right state $q_r$, $\lambda$ must increase. In Figure \ref{fig:PW_curves} (c) and (d) we see the Hugoniot loci and integral curves for this system with the contour lines for $\lambda_1$ and $\lambda_2$. Using these plots, we can determine which areas of the curves are valid shocks and which are valid rarefaction waves. Figure \ref{fig:PW_curves} (e) and (d) shows the points $q_l$ and $q_r$, respectively, with only the valid waves displayed. For these plots,
$q_* = (\rho_*,\rho_*v_*) = (0.2, 0.04)$ and $\gamma = 0.2$. 
Here, shocks are shown in red or magenta and rarefaction waves are shown in either light or dark green.

@@

In conservative form, the second equation becomes $\Big( \rho (v + p(\rho)) \Big)_t + \Big( \rho v (v + p(\rho)) \Big)_x = 0$.  Aw and Rascle call this conserved quantity, $y = \rho v + \rho p(\rho)$,``momentum", although they admit that $y$ has no obvious physical interpretation. Furthermore, the second equation can also be rewritten as an equation for velocity: $v_t + \Big(v - \rho p'(\rho)\Big)v_x = 0$. If we consider the linearized system,

@@

\subsubsection{Hugoniot Loci}
In order to consider the Hugoniot loci for this system we must first write the system in conservation form. 
To accomplish this, note that multiplying the first equation by $(v + p(\rho ))$ and the second equation 
by $\rho$ gives us the two equations 
\begin{align*}
&(v + p(\rho ))\partial_t\rho + (v + p(\rho ))\partial_x(\rho v) = 0,  \\
&\rho\partial_t \left(v + p(\rho )\right) + \rho v\partial_x \left( v + p(\rho )\right) = 0.
\end{align*}
Adding these two equations gives
\begin{align}\label{AR:eq2}
\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
\end{align}
Then from (\ref{AR:eq1}) and (\ref{AR:eq2}) we have the system 
\begin{align*}
&\partial_t\rho + \partial_x(\rho v) = 0, \\
&\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
\end{align*}
This can be rewritten as $q_t + f(q)_x = 0$ by defining
\begin{align*}
q = \left[ \begin{matrix}
\rho \\ y
\end{matrix}\right], \hspace{0.3in}
f(q) = \left[ \begin{matrix}
v\rho \\
vy
\end{matrix}\right].
\end{align*}
where $y = \rho\left(v + p(\rho )\right)$.
The Rankine-Hugoniot condition tells us that 
\begin{align*}
s(q_*- q) = f(q_*) - f(q).
\end{align*}
Therefore, from this condition we get the two equations
\begin{align}
s(\rho_* - \rho) = v_*\rho_* - v\rho, \label{AR:eq3}\\
s(y_* - y) = v_*y_* - vy.\label{AR:eq4}
\end{align}
With a little rearranging and by plugging in for $y$, these two equations can be rewritten as
\begin{align}
&\rho_*(s - v_*) = \rho (s - v), \label{AR:eq5}\\
&\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) = \rho(s - v)\left(v + p(\rho )\right).\label{AR:eq6}
\end{align}
From equation (\ref{AR:eq5}), we know that there are two cases:

(a) $\rho_*(s - v_*) = \rho (s - v) = 0$. In this case we have that
\begin{align*}
s - v_* = s - v 
\end{align*}
except if one of the two densities, $\rho_*$, $\rho$ is zero. Therefore, we can say that
\begin{align*}
v_* = v.
\end{align*}
Since the velocity does not change, this is a contact discontinuity. Using the fact that $v = v_*$ in (\ref{AR:eq4}) gives
\begin{align*}
y_*(s - v) = y(s - v).
\end{align*}
So, either $y = y_*$ or $s = v$. If $y = y_*$ we have a stationary point in the $Y = (\rho, y)$ plane, which does not 
describe a wave. Therefore, $s = v$. Note that we have no restrictions on $y$ other than its definition. 
Therefore this loci in the $Y = (\rho, y)$ plane is given by 
\begin{align*}
y = \rho ( v_* + p(\rho )).
\end{align*}

@@

(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into equation (\ref{AR:eq6}) gives
\begin{align*}
\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
\rho(s - v)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
v_* + p(\rho_* )&= v + p(\rho ).
\end{align*}
Therefore, the Hugoniot loci for this case in the $Y = (\rho, y)$ plane are given by
\begin{align*}
y = v_* + p(\rho_* ),
\end{align*}
since $y = v + p(\rho )$ by definition. Note that solving equation (\ref{AR:eq3}) for $s$ gives
\begin{align*}
s = \frac{\rho_*v_* - \rho v}{\rho_* - \rho}.
\end{align*}
Using the fact that $v_* = v + p(\rho) - p(\rho_*)$, we get that
\begin{align*}
s = \frac{\rho_*\left( v + p(\rho) - p(\rho_*)\right)- \rho v}{\rho_* - \rho}
= v - \rho_*\frac{p(\rho ) - p(\rho_*)}{\rho - \rho_*}
\approx v - \rho p'(\rho)
\end{align*}
for $\rho_* \approx \rho$.

In the $U = (\rho, v)$ plane these loci become (a) $v = v_*$, $\rho$ free, and (b) $v = v_* + p(\rho_* ) - p(\rho )$.
 In the $M = (\rho, \rho v)$ plane these loci become (a) $\rho v = \rho_* v_*$, $\rho$ free, 
 and (b) $\rho v = \rho \left( v_* + p(\rho_*)\right) - \rho p(\rho))$. Figure \ref{fig:AR_curves} shows the Hugoniot loci 
 in the M plane. Note that for these figures we used $\gamma = 0.2$, and in Figure \ref{fig:AR_curves} (c) we used $\rho_* = 0.2$ and $m_* = 0.04$.

\subsubsection{Integral Curves}
In order to consider the integral curves for this system \cite{AwRascle2000} 
works from the system
\begin{align*}
&\partial_t \rho + \partial_x (\rho v) = 0 \\ 
&\partial_t v + \left(v - p'(\rho)\rho\right)\partial_x v = 0,
\end{align*}
where the second equation is found by multiplying equation (\ref{AR:eq1}) 
by $p'(\rho)$ and subtracting it from equation (\ref{AR:eq1.5}). 
This can be rewritten as $q_t + Aq_t = 0$ by defining
\begin{align*}
q = \left[ \begin{matrix}
\rho \\ v
\end{matrix}\right] , \hspace{0.3in}
A = \left[ \begin{matrix}
v & \rho \\
0 & v - p'(\rho ) \rho
\end{matrix}\right].
\end{align*}
The eigenvalues of $A$ are $\lambda_1 = v - p'(\rho ) \rho$ and 
$\lambda_2 = v$. Note that since $p(\rho ) = \rho^{\gamma}$ 
and $\rho \geq 0$, $\lambda_1 < \lambda_2$ so long as $\rho \neq 0$. 
Also note that if $\rho = 0$ then $A$ is no longer 
hyperbolic. The eigenvectors of $A$ are
\begin{align*}
r_1 = \left[ \begin{matrix}
1 \\ - p'(\rho )
\end{matrix}\right], \hspace{0.3in}
r_2 = \left[ \begin{matrix}
1 \\ 0
\end{matrix}\right].
\end{align*}
Using equation (13.24) from \cite{LeVeque2002} with $\alpha = 1$ tells us 
that the integral curves can be found by considering $q'(\xi ) = r_p$. 
For $p = 1$ this gives us the two equations
\begin{align*}
&\rho'(\xi ) = 1 \\
&v'(\xi ) = - p'(\rho ),
\end{align*}
which can be solved to find that we have parameratized our curve by $\rho$, 
and $v = - p'(\rho ) + c_1$ while $\rho$ is free, where $c_1$ is a constant. 
Since for a given point $(\rho_*, v_*)$ when $\rho = \rho_*$ we want 
$v = v_*$, we get that $c_1 = v_* + p'(\rho_*)$. Thus, the integral curve 
is given by 
\begin{align*}
v = v_* + p'(\rho_*) - p'(\rho ).
\end{align*}
This coincides with the Hugoniot loci from case (b). For $p = 2$, we get the 
two equations
\begin{align*}
&\rho'(\xi ) = 1 \\
&v'(\xi ) = 0,
\end{align*}
which can be solved to find that we have parameratized our curve by $\rho$, 
and that $v$ is a constant. Therefore, $v = v_*$ and $\rho$ is free. 
This coincides with the Hugoniot loci from case (a).

\begin{figure}[H]
 \centering
    \subfigure[$\lambda_1$-curves with contours of $\lambda_1$.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M.jpg}
   }
 \subfigure[Valid waves for the point $q_l$.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_ql.jpg}
   }
 \subfigure[Valid waves for the point $q_r$.]{
  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_qr.jpg}
   }
 \caption[Optional caption for list of figures]
 {Figures showing the calculation of the valid waves for a given point. Figure (a)
  shows the $\lambda_1$ curves in black with the $\lambda_1$ contours 
  in cyan. The other figures show valid shocks in red and valid rarefaction 
  waves in green. The contact discontinuity is shown in blue.}
   \label{fig:AR_validity}
\end{figure}

@@

Therefore, it is important to take into account the contour lines of $\lambda$ when 
looking at the loci and integral curves. Figure \ref{fig:AR_validity}(a) 
shows the $\lambda_1$-curves in the $M$ plane,
with the contour lines of $\lambda_1$ shown in magenta. This information is used
to determine which areas of the $\lambda_1$ curves are valid shock waves and which
are valid rarefaction waves. The rest of the subfigures show the regions of validity for 
specific points $q_*$, depending on whether $q_* = q_l$ or $q_* = q_r$. For these plots,
$q_* = (\rho_*,\rho_*v_*) = (0.2, 0.04)$ and $\gamma = 0.2$. Note that the contact discontinuity is valid
anywhere in the plane since $\lambda_2$ is linearly degenerate.

@@

@@

Reproducing the Riemann problems found in \cite{AwRascle2000}, we consider two different examples. The variables used in these two examples are shown in Table \ref{table:1}.

In both examples the density and velocity are greater in the right state than in the left state.  This corresponds to the situation considered by Aw and Rascle that motivated the use of the convective derivative.  In the first example, we see that the AR model predicts that the density of the middle state is less than both left or right states, but that the velocity increases steadily from left state to the middle state and then remains constant while the density changes due to the contact discontinuity.  This is what we would expect, as drivers spread out as they accelerate to catch up with the cars ahead.  On the other hand, the PW model predicts that the middle state will have a greater density than the left state but a smaller velocity.  This is opposite of what we would expect.

@@

\begin{figure}[H]
 \centering
 \subfigure[Solution from the AR model. The transition from $q_l$ to $q_r$ is first a shock and then a contact discontinuity.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example1.jpg}
   }
    \subfigure[Solution from the PW model.  The transition from $q_l$ to $q_r$ is first a shock and then a rarefaction wave. ]{
  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example1.jpg}
   }
 \caption[Optional caption for list of figures]
 {Solutions to the Riemann problem for example 1 from the AR and PW models. The middle state is denoted by $q_m$.}
  \label{fig:example1}
\end{figure}

\begin{figure}[H]
 \centering
 \subfigure[Solution from the AR model. The transition from $q_l$ to $q_r$ is a contact discontinuity,  shown in blue, there is no middle state.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example2.jpg}
   }
 \subfigure[Solution from the PW model. The transition from $q_l$ to $q_r$ is a rarefaction wave, shown in dark green, there is no middle state.]{
  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example2.jpg}
   }
 \caption[Optional caption for list of figures]
 {Solutions to the Riemann problem for example 2 from the AR and PW models.}
  \label{fig:example2}
\end{figure}
